---
title: "Moneyball: DATA621 Project #1"
author: "Team 2. "
date: "03/07/2021"
output: 
  pdf_document:
    toc: true
    number_sections: true
    highlight: tango
    df_print: kable
fontsize: 11pt
geometry: margin=1in
---

File to submit.

## Introduction

[add text intro here]

## Setup

This analysis requires installation of `tidyverse`,`corrplot`, and `reshape2` [add other pckgs gere]

```{r, setup, include=FALSE} 
knitr::opts_chunk$set(fig.width = 10, message=FALSE, warning=FALSE, cache = TRUE)
#This will not appear in the PDF. 
#This setting hides all the warnings and messages in the knit document (not in the R!), and caches the results of heavy chunks so it loads faster.
```

```{r pckgs, include=FALSE}
#This will not appear in the PDF. 
library(dplyr)
library(tidyverse)
library(reshape2)
library(corrplot)
```

### Load Raw Data

The data has been divided in advance into a training set, `training_raw`, and an evaluation set, `eval_raw`.  The evaluation set does *not* contain the target information. 

```{r load_data, message=FALSE}
url1 = "https://raw.githubusercontent.com/Jagdish16/CUNY_DATA_621/main/project_1/moneyball-training-data.csv"
training_raw = read_csv(url1)
```

### Rename Columns

Rename the columns to be more human-readable.

```{r rename}
training <- training_raw %>%
  select(-INDEX) %>%
  rename_with(~ gsub("TEAM_", "", .x)) %>%
  rename_with(stringr::str_to_title) %>%
  dplyr::rename(
    Wins = Target_wins,
    Hits = Batting_h,
    Doubles = Batting_2b,
    Triples = Batting_3b,
    HomeRuns = Batting_hr,
    Walks_AtBat = Batting_bb,
    StrikeOuts_AtBat = Batting_so,
    BasesStolen = Baserun_sb,
    OutStealingBases = Baserun_cs,
    Hits_Allowed = Pitching_h,
    HitByPitch_AtBat = Batting_hbp,
    Errors = Fielding_e,
    HomeRuns_Allowed = Pitching_hr,
    Walks_Allowed = Pitching_bb,
    StrikeOuts = Pitching_so,
    DoublePlays = Fielding_dp
  )
```


## Data Exploration

What does the data look like? 

``` {r boxplots, warning=FALSE, message=FALSE}
long <- training %>% as.data.frame() %>% melt()

long %>%
  ggplot(aes(x=value)) + 
  geom_boxplot() + 
  facet_wrap(~variable, scales='free')
  
```

### Missing cases

Several variables have missing values. HitByPitch_AtBat has 1678 missing!

By default, R's built-in `lm` function will omit cases with missing values.

```{r missing}
missing <- sapply(training, FUN = function(x) sum(is.na(x)))
missing
```


### Feature Correlation

```{r}
# computing correlation matrix
corr <- round(cor(training), digits = 1)

corrplot(corr, type = 'upper', addCoef.col = 'black', tl.col = 'black', tl.srt = 45)
```


## Models

### Handling of NA values

Having tested both imputing value for missing variables and omitting cases with missing variables, a better result is found for omitting cases. [Question -- could this be because we eliminated too many cases by dropping the cases with missing at bats?]

#### All features

#### All features with imputed values for NAs

### Transform data
#### Box-Cox Transform

### Feature selection

[select out features with lowest correlation]

## Testing Model Assumptions

[residual plots]

## Model Selection & Conclusions

Choose model with most explanatory value (high adjusted $R^2$) with fewest features. Assume that the simplest model should be preferred, all else being equal.

## Prediction

The final exercise is to use the selected model to predict the target values (wins) for the eval set, `eval_raw`.

```{r predict}
url2 = "https://raw.githubusercontent.com/Jagdish16/CUNY_DATA_621/main/project_1/moneyball-evaluation-data.csv"
eval_raw=read_csv(url2)
```

## References

Sellmair, Reinhard. "How to handle correlated features?" June 25, 2018. https://www.kaggle.com/reisel/how-to-handle-correlated-features

Xie, Yihui, J. J. Allaire, and Garrett Grolemund, *R Markdown: The Definitive Guide*, CRC PressDecember 14, 2020 https://bookdown.org/yihui/rmarkdown/r-code.html. 
