---
title: "DS621_HW1_Jagdish"
author: "Jagdish Chhabria"
date: "2/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(MASS)
library(tidyverse)
library(dplyr)
library(reshape2)
library(kableExtra)
library(corrplot)
library(ggplot2)
library(Hmisc)
library(PerformanceAnalytics)
library(GGally)
library(ggpubr)
library(car)

```


## R Markdown

```{r }
Load data
url = "https://raw.githubusercontent.com/Jagdish16/CUNY_DATA_621/main/project_1/moneyball-training-data.csv"
train.raw = read.csv(url)
#train.raw<-read.csv("moneyball-training-data.csv")
train.raw<-subset(train.raw,select=-c(INDEX))

```



```{r}
# Inspect data
dim(train.raw)
summary(train.raw)
```

```{r}
# Remove selected columns with large number of missing values
train.raw<-subset(train.raw,select=-c(TEAM_BATTING_HBP,TEAM_BASERUN_CS))

```

```{r}
# Check updated data frame
summary(train.raw)
```

The following columns look like they have significant outliers:
- TEAM_PITCHING_BB
- TEAM_BASERUN_SB
- TEAM_PITCHING_SO

```{r}
# Check columns that seem to have outliers

#ggplot(stack(train),aes(x=ind,y=values))+geom_boxplot()
#ggplot(melt(train), aes(variable, value)) + geom_boxplot()
boxplot(train.raw$TEAM_PITCHING_SO)
boxplot(train.raw$TEAM_PITCHING_BB)
boxplot(train.raw$TEAM_BASERUN_SB)

```


```{r}
# Remove outliers - Method 1
for (n in c("TEAM_PITCHING_BB", "TEAM_BASERUN_SB", "TEAM_PITCHING_SO")) {
  outliers<-boxplot(train.raw[,n],plot=FALSE)$out
  train_updated<-train.raw[-which(train.raw[,n] %in% outliers),]
}

```


```{r}
summary(train_updated)
```



```{r}
# Check impact of removing outliers
boxplot(train_updated$TEAM_PITCHING_SO)
boxplot(train_updated$TEAM_PITCHING_BB)
boxplot(train_updated$TEAM_BASERUN_SB)

```

```{r}
# Impute missing values with the median value for each remaining column

train<-data.frame(sapply(train_updated,function(x) ifelse(is.na(x),median(x,na.rm = TRUE),x)))

```



```{r}
# Check summary of modified dataset
summary(train)

```




```{r data Distributions, fig.height = 10, fig.width = 10, error=FALSE, warning=FALSE}

# Check distribution of remaining predictors
distributions<-train%>%gather(key='variable_name', value='value')
ggplot(distributions)+geom_histogram(aes(x = value, y = ..density..), bins = 30)+geom_density(aes(x = value), color = 'red')+facet_wrap(. ~ variable_name, scales = 'free', ncol = 4)

```

```{r}
# Calculate correlation between variables
train_corr_matrix<-train%>%cor()%>%round(2)%>%as.matrix()
train_corr_matrix%>%kable()%>%kable_styling()

```


```{r}
# flattenCorrMatrix
# cormat : matrix of the correlation coefficients
# pmat : matrix of the correlation p-values
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
```


```{r}
# Another method to check correlations and their significance
#train%>%pairs()
#train%>%ggpairs()
corr.mat<-rcorr(as.matrix(train))
#corr.mat

flattenCorrMatrix(corr.mat$r,corr.mat$P)%>%arrange(row,-abs(cor))

```

We drop the following columns:
- TEAM_BATTING_HBP
- TEAM_PITCHING_H
- TEAM_PITCHING_SO

We transform the following columns:
- TEAM_BASERUN_SB
- TEAM_BASERUN_CS
- TEAM_BATTING_3B
- TEAM_BATTING_HR
- TEAM_BATTING_SO
- TEAM_FIELDING_DP
- TEAM_FIELDING_E
- TEAM_PITCHING_HR


```{r}
par(mfrow=c(3,5))

for (i in 1:ncol(train[,1: ncol(train) - 1 ])) {  
  qqnorm(train[,i], main=names(train[i]))
  qqline(train[,i])
}

```


```{r}
# Rename columns
train<-train%>%rename(wins=TARGET_WINS, baserun_sb=TEAM_BASERUN_SB, bat_3b=TEAM_BATTING_3B, bat_hr=TEAM_BATTING_HR, bat_so=TEAM_BATTING_SO, field_dp=TEAM_FIELDING_DP, field_e=TEAM_FIELDING_E, pitch_hr=TEAM_PITCHING_HR, pitch_h=TEAM_PITCHING_H, pitch_so=TEAM_PITCHING_SO, bat_h=TEAM_BATTING_H, bat_2b=TEAM_BATTING_2B, bat_bb=TEAM_BATTING_BB, pitch_bb=TEAM_PITCHING_BB)

```


```{r}
head(train)
```


```{r}
#fit linear regression model
model<-lm(wins~bat_h+bat_2b+bat_bb+bat_3b+bat_hr+pitch_bb+pitch_hr, data=train)

#find optimal lambda for Box-Cox transformation 
bc<-boxcox(wins~bat_3b, data=train)
(lambda<-bc$x[which.max(bc$y)])

#fit new linear regression model using the Box-Cox transformation
new_model<-lm(((wins^lambda-1)/lambda)~bat_3b, data=train)

```


```{r}
# Validate both models
#define plotting area
op <- par(pty = "s", mfrow = c(1, 2))

#Q-Q plot for original model
qqnorm(model$residuals)
qqline(model$residuals)

#Q-Q plot for Box-Cox transformed model
qqnorm(new_model$residuals)
qqline(new_model$residuals)

#display both Q-Q plots
par(op)



```






```{r}
# Fit a model with selected variables
mdl1<-lm(wins~bat_h+bat_2b+bat_bb+bat_3b+bat_hr+pitch_bb+pitch_hr, data=train)

summary(mdl1)

```



