---
title: "project 1"
author: "Orli Khaimova"
date: "2/24/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(corrplot)
library(ggplot2)
library(tidyverse)
library(magrittr)
library(Hmisc)
library(e1071)
```

```{r}
moneyball <- read.csv("https://raw.githubusercontent.com/Jagdish16/CUNY_DATA_621/main/project_1/moneyball-training-data.csv") 

```


### Data Exploration

The following variables have missing data:
* TEAM_BATTING_HBP, 2085
* TEAM_BASERUN_CS, 772
* TEAM_FIELDING_DP, 286
* TEAM_BASERUN_SB, 131
* TEAM_BATTING_SO, 102
* TEAM_PITCHING_SO, 102

```{r}
summary(moneyball)

# creates a list with the correlations and p values 
corr_data<- rcorr(as.matrix(moneyball))
corr_p <- round(corr_data$P,4)

# this takes the values and correlations and makes it into a 2 column dataframe
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}

# sorted the pairs of correlations by their p value to show variables with the biggest
# relationships, with the p showing the significance value

sorted_cor <- flattenCorrMatrix(corr_data$r, corr_data$P) %>%
   arrange(desc(abs(cor)))

# As seen, TEAM_BATTING_HR is correlated to TEAM_PITCHING_HR and TEAM_BATTING_SO

head(sorted_cor) #this shows the highest correlations

# correlation plot
# didn't remove the index yet
corrplot(corr_data$r, type = "upper", order = "hclust", tl.col = "black")

# prints out the box plots for each variable
moneyball %>%
  dplyr::select(-c("INDEX")) %>%
  gather(variable, value, TARGET_WINS:TEAM_FIELDING_DP) %>%
  ggplot(., aes(x= variable, y=value)) + 
  geom_boxplot() +
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = element_blank())

# histograms for each variable
moneyball %>%
  dplyr::select(-c("INDEX")) %>%
  gather(variable, value, TARGET_WINS:TEAM_FIELDING_DP) %>%
  ggplot(., aes(x = value)) + 
  geom_histogram(bins = 25) +
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = element_blank())


# this is to just show the skewness of each variable
# TEAM_PITCHING_SO, TEAM_PITCHING_H, TEAM_PITCHING_BB are highly skewed
# can fix skewness with transformation such as log or inverse

moneyball %>%
  dplyr::select(-c("INDEX")) %>%
  sapply(., skewness, na.rm = TRUE) %>% 
  as.data.frame() %>%
  arrange(desc(.))

```

### Data Preparation

* Created TEAM_BATTING_1B, singles by batters
* removed TEAM_BATTING_HBP, imputing with mean/mode gave high P value

```{r}
moneyball <- moneyball %>% 
  mutate(TEAM_BATTING_1B = TEAM_BATTING_H - TEAM_BATTING_2B - TEAM_BATTING_3B - TEAM_BATTING_HR) %>%
  dplyr::select(-c("TEAM_BATTING_HBP"))

# -----------------------------------------------------------------------------  

#can ignore, I tried to transform the variable with big skewness
#but it did not affect the model
#could be because the model is removing the NA's

# test <- moneyball %>%
#  mutate(TEAM_PITCHING_SO = ifelse(is.na(TEAM_PITCHING_SO), 
#                                 mean(moneyball$TEAM_PITCHING_SO, na.rm = TRUE),TEAM_PITCHING_SO),
#         TEAM_PITCHING_SO = 1/TEAM_PITCHING_SO,
#         TEAM_PITCHING_SO = ifelse(is.infinite(TEAM_PITCHING_SO),NA,TEAM_PITCHING_SO)) %>%
#  dplyr::select(TEAM_PITCHING_SO)

# skewness(test$TEAM_PITCHING_SO, na.rm = TRUE)

# went from 22 to 1.68

#moneyball$TEAM_PITCHING_SO <- test$TEAM_PITCHING_SO

# -----------------------------------------------------------------------------  

#ignore, in case of imputing
#TEAM_BATTING_HBP = ifelse(is.na(TEAM_BATTING_HBP), mean(moneyball$TEAM_BATTING_HBP, na.rm = TRUE),TEAM_BATTING_HBP)

# can ignore this
# this was to show the difference between the two
difference <- moneyball %>%
  dplyr::select(c("TEAM_PITCHING_HR","TEAM_BATTING_HR")) %>%
  mutate(difference = TEAM_PITCHING_HR-TEAM_BATTING_HR)
summary(difference)
```

### Build Models


```{r}
#highest model was with team_batting_hbp included, prior to removing any na

#Removed: TEAM_BATTING_H (no diff to model)


# I applied logs and inverse to the variables to try to fix the skewness
# but R2 stayed the same, I believe due to NA


model1 <- lm(TARGET_WINS ~   TEAM_BATTING_2B + TEAM_BATTING_3B + 
              TEAM_BATTING_HR + TEAM_BATTING_BB  + TEAM_BATTING_SO +TEAM_BASERUN_SB +  
              TEAM_BASERUN_CS + TEAM_PITCHING_H + TEAM_PITCHING_HR +  
              TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP +
              TEAM_BATTING_1B, moneyball)

summary(model1)

plot(fitted(model1), resid(model1), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)
qqnorm(resid(model1), main = "Normal Q-Q Plot", col = "darkgrey")
qqline(resid(model1), col = "dodgerblue", lwd = 2)
```

```{r}
#took log, .4295
#removed TEAM_BATTING_SB, .4299
#removed TEAM_PITCHING_HR, .4302
# removed TEAM_BATTING_SO, .4305
# removed based on P values
model2 <- lm(log(TARGET_WINS) ~   TEAM_BATTING_2B + TEAM_BATTING_3B + 
              TEAM_BATTING_HR + TEAM_BATTING_BB +TEAM_BASERUN_SB +  
              TEAM_BASERUN_CS + TEAM_PITCHING_H + 
              TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP +
              TEAM_BATTING_1B, moneyball)

summary(model2)

plot(fitted(model2), resid(model2), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)
qqnorm(resid(model2), main = "Normal Q-Q Plot", col = "darkgrey")
qqline(resid(model2), col = "dodgerblue", lwd = 2)
```

```{r}
#boxcox method

model3 <- lm(TARGET_WINS ~   TEAM_BATTING_2B + TEAM_BATTING_3B + 
              TEAM_BATTING_HR + TEAM_BATTING_BB  +TEAM_BASERUN_SB +  
              TEAM_BASERUN_CS + TEAM_PITCHING_H + 
              TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP +
              TEAM_BATTING_1B, moneyball)
summary(model3)

MASS::boxcox(model3, lambda = seq(0.75, 1.6, by = 0.05), plotit = TRUE)

model_cox = lm((((TARGET_WINS ^ 1.23) - 1) / 1.23) ~ TEAM_BATTING_2B + TEAM_BATTING_3B + 
              TEAM_BATTING_HR + TEAM_BATTING_BB  +TEAM_BASERUN_SB +  
              TEAM_BASERUN_CS + TEAM_PITCHING_H + 
              TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP +
              TEAM_BATTING_1B, moneyball)

summary(model_cox)

plot(fitted(model_cox), resid(model_cox), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)
qqnorm(resid(model_cox), main = "Normal Q-Q Plot", col = "darkgrey")
qqline(resid(model_cox), col = "dodgerblue", lwd = 2)
```

```{r}
#took ratio of the correlated variables

model4 <- lm(TARGET_WINS ~   TEAM_BATTING_2B + TEAM_BATTING_3B + 
              I( TEAM_PITCHING_HR / TEAM_BATTING_HR) +  
              TEAM_BASERUN_CS + TEAM_PITCHING_H + TEAM_BASERUN_SB +
              TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP +
              TEAM_BATTING_1B, moneyball)
summary(model4)

plot(fitted(model4), resid(model4), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)
qqnorm(resid(model4), main = "Normal Q-Q Plot", col = "darkgrey")
qqline(resid(model4), col = "dodgerblue", lwd = 2)
```


```{r}
#here applies random transformations
model5 <- lm(TARGET_WINS ~ TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B + 
               I(TEAM_PITCHING_HR*TEAM_BATTING_HR) + TEAM_BATTING_BB + TEAM_FIELDING_E +
               TEAM_BASERUN_SB + I(1/TEAM_FIELDING_DP) + TEAM_PITCHING_SO +
               I(TEAM_BASERUN_CS ^ 3), moneyball)


summary(model5)

plot(fitted(model5), resid(model5), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)
qqnorm(resid(model5), main = "Normal Q-Q Plot", col = "darkgrey")
qqline(resid(model5), col = "dodgerblue", lwd = 2)
